Model-based pretreatment for regression with correlated variables.

### 2014-04-08 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@507 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
	\begin{eqnarray}
		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \\
		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)\\  
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.

@@

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$. The $\delta_j$ and $\beta_j$ are generated according to the same Poisson law but with a random sign. $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.  
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.

		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure?}
			The first criterion is $\psi$ which is minimised in the MCMC. But in our case, $\psi$ is based on the likelihood whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
			Global indicators :
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_2\cap \hat{I}_2|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_2|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_2|-TL$
				\item $\Delta p_2$ : the gap between the number of sub-regression in both model : $\Delta p_2=|I_2|-|\hat{I}_2|$. The sign defines if $\hat{S}$ is too complex or too simple
				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_2}p_1^j-\sum_{j \in \hat{p}_2}\hat{p}_1^j$
			\end{itemize}
		\subsubsection{Results on $S$}	\label{compZ}
In table \ref{compZvrai} we compare found structures in different contexts with both Uniform (U) and Hierarchical Uniform (HU) a priori law on $P(S)$. 
We see that (HU) hypothesis gives sparser models even if $\max (\hat{p}_1^j)=5$. $p=40$ so the maximum value for $\hat{p_2}$ is $20$ in (HU) case and we see that this max is not reached. The stronger penalty implied by (HU) really is efficient. The datasets used for (U) and (HU) are the same to keep the comparison  meaningful. These datasets and $\hat{S}$ are those used for tables 
 \ref{YXlinOLS} to \ref{YX2linstep}.

 It is also notable that (HU) has a greater computational cost than (U). But next versions of the package may optimize a bit more the penalization step and reduce this gap.

@@

$n$ & $p_2$ & $\psi$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\

@@

\clearpage
\subsection{Results on prediction}
	\subsubsection{$Y$ depends on all variables in $X$}	 \label{tableMSEsimtout}	
We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure). The datasets used here were those from \ref{compZvrai}.

We observe that \textsc{CorReg} is better than classical methods especially with the Hierarchical Uniform law. When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods. We also observe in table \ref{YXlinstep} that simple methods like stepwise (here from the package \textsc{lars}) can give good results in prediction.

When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}).

>>>>

	We note $\boldsymbol{\Theta}$ the set of the parameters of the generative model
	\begin{eqnarray}
		P(S|\boldsymbol{X})&\propto& P(\boldsymbol{X}|S)P(S) \label{approxBIC}\\
		-2\log P(\boldsymbol{X}|S)&\approx & BIC=-2\mathcal{L}(\boldsymbol{X},S,\boldsymbol{\Theta})+|\boldsymbol{\Theta}|\log(n)\\  
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
		We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.

@@

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, $\sigma_Y=10$, $\sigma=0.001$, variables $X_1$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda$ and which standard deviation also is $\lambda$. The $\boldsymbol{\alpha_j}$ and $\beta_j$ are generated according to the same Poisson law but with a random sign. $S$ only contains binary relationships but \textsc{CorReg} was only constrained to $\max (\hat{p}_1^j)=5$.  
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the walk was launched on $20$ initial structures with a maximum of 9 000 steps each time.

		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure?}
			The first criterion is $P(S|\boldsymbol{X})$ which is maximized in the MCMC. But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
			Global indicators :
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_2\cap \hat{I}_2|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_2|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_2|-TL$
				\item $\Delta p_2$ : the gap between the number of sub-regression in both model : $\Delta p_2=|I_2|-|\hat{I}_2|$. The sign defines if $\hat{S}$ is too complex or too simple
				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_2}p_1^j-\sum_{j \in \hat{p}_2}\hat{p}_1^j$
			\end{itemize}
		\subsubsection{Results on $S$}	\label{compZ}
In table \ref{compZvrai} we compare found structures in different contexts with both Uniform (U) and Hierarchical Uniform (HU) a priori law on $P(S)$. 
We see that (HU) hypothesis gives sparser models even if $\max (\hat{p}_1^j)=5$. $p=40$ so the maximum value for $\hat{p_2}$ is $20$ in (HU) case and we see that this max is not reached, meaning that the stronger penalty implied by (HU) really is efficient. The datasets used for (U) and (HU) are the same to keep the comparison  meaningful. These datasets and $\hat{S}$ are those used for tables 
 \ref{YXlinOLS} to \ref{YX2linstep}.

 It is also notable that (HU) has a greater computational cost than (U). 

@@

\multicolumn{3}{|c}{configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
$n$ & $p_2$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_2$ & $\Delta compl$ \\

@@

\clearpage
\subsection{Results on prediction}
	\subsubsection{$Y$ depends on all variables in $X$}	 \label{tableMSEsimtout}	
We first try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure). The datasets used here were those from table \ref{compZvrai}.

We observe that \textsc{CorReg} is better than classical methods especially with the Hierarchical Uniform law. When the complexity of the true model is higher than $\frac{p}{2}$ Uniform hypothesis logically is better but (HU) still beats classical methods, so it is robust. We also observe in table \ref{YXlinstep} that simple methods like stepwise (here from the package \textsc{lars}) can give good results in prediction.

When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
